|

### 2020-05-19 paper/Manuscript.Rmd

author:
  - name: "Caspar J. Van Lissa"
    institution: "1,2"
    orcid: "0000-0002-0808-5024"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    institution: "3,4"
    orcid: "0000-0001-8765-6982"
  - name: "Loek Brinkman"
    institution: "2,5"
    orcid: "0000-0003-3997-1173"
  - name: "Aaron Peikert"
    institution: "3,6"
    orcid: "0000-0001-7813-818X"
  - name: "Barbara M.I. Vreede"
    institution: "7"
    orcid: "0000-0002-5023-4601"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "UMC Utrecht"
  - id: "6"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "7"
    institution: "University Library Utrecht"

@@

The TOP-guidelines are primarily geared towards top-down systemic change by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement.
This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions, and where such requirements do exist, this workflow will help fulfill them.

@@

## Prerequisites

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine that can be stored like a sort of time capsule, and identically reinstated on different systems. This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data). <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.
<!-- Barbara: I see no difference between these two symbols...-->

@@

* `open_data(data)`: This function stores the `data` object (a `data.frame` or `matrix`) in a spreadsheet file (called `data.csv` by default).
This file should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
Once the GitHub repository is made public, these data will be open to the public.
Assume that, once you do this, the data cannot be un-shared.
* `closed_data(data)`: This function also stores the data in a local file (called `data.csv` by default), but that file is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
The function also adds a checksum for the original data in the `.worcs` project file, and creates a synthetic dataset (called `synthetic_data.csv` by default).
The user should commit all changed files and push them to GitHub.
Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.

>>>>

author:
  - name: "Caspar J. Van Lissa"
    institution: "1,2"
    orcid: "0000-0002-0808-5024"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    institution: "3,4"
    orcid: "0000-0001-8765-6982"
  - name: "Loek Brinkman"
    institution: "2,5"
    orcid: "0000-0003-3997-1173"
  - name: "Aaron Peikert"
    institution: "3,6"
    orcid: "0000-0001-7813-818X"
  - name: "Barbara M.I. Vreede"
    institution: "7"
  - name: "Anna-Lena Lamprecht"
    institution: "2,8"
    orcid: "0000-0002-5023-4601"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "UMC Utrecht"
  - id: "6"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "7"
    institution: "University Library Utrecht"
  - id: "8"
    institution: "Utrecht University, Department of Information and Computing Sciences"

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement [REF Loek paper].
This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions:
If such requirements exist, this workflow will help fulfill them.

@@

<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers. For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more. I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better. --> 
## Prerequisites

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, one might use the
aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all
analyses inside a "container": An environment that behaves like a virtual computer that can be
stored like a sort of time capsule, and identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g., preprocessing tools
for neuroimaging data). However, it can be difficult to set up for novice users - although a package
is in development to facilitate this transition [@repo]. The cloud-based research collaboration platform
'Code Ocean' is also based on Docker, for example. It is user-friendly, but it is not clear whether the
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/) is fully
GDPR compliant - it seems to focus more on limiting liability than on compliance. This is a risk with
any cloud-based solution. The advantage of using a local system is that data sharing is optional, and
WORCS capitalizes on this flexibility.

A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software
used in a project to a dedicated library that can be stored and copied with the project. This solution
is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages
must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address
the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is
somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing
all software used, it simply records the calendar date, and then installs the version of all required
packages that was available on the central R repository `CRAN` on that date from an archival website.
Thus, all that is needed to restore a project is the date on which it was conducted. This is very
user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN`
packages, and cannot mix package versions from different dates.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going
back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
Users can render a manuscript to PDF either with, or without, comprehensive citations.

@@

* `open_data(data)`: This function stores the `data` object (a `data.frame` or `matrix`) in a spreadsheet file (called `data.csv` by default).
This file should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
Once the GitHub repository is made public, these data will be open to the public.
Assume that, once you do this, the data cannot be un-shared.
* `closed_data(data)`: This function also stores the data in a local file (called `data.csv` by default), but that file is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
The function also adds a checksum for the original data in the `.worcs` project file, and creates a synthetic dataset (called `synthetic_data.csv` by default).
The user should commit all changed files and push them to GitHub.
Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->
